A meta-analysis of powdery mildew yield loss mitigation strategies in Australian mungbean

### 2021-03-10 03_Yield_MetaAnalysis.Rmd
Clarification and cleaning in analysis

```{r Yield_ME_libraries, echo=TRUE,results='hide'}
if (!require("pacman"))
   install.packages("pacman")
pacman::p_load(tidyverse,
               kableExtra,
               bomrang,
               RColorBrewer,
               metafor,
               here,
               netmeta,
               multcomp,
               flextable)

@@

# Data
PM_dat_Y <-
   read.csv("cache/PM_yield_clean_data.csv", stringsAsFactors = FALSE)
```

@@

## Spray schedule meta-analysis  
### Spray schedule moderator

@@

```{r Yield_Metafor-analysis}
PM_mv <- rma.mv(
   yi = grain_yield.t.ha.,
   vi,
   mods = ~ spray_management,
   method = "ML",
   random = list(~ spray_management | trial, ~ 1 | id),
   struct = "UN",
   control = list(optimizer = "optim"),
   data = PM_dat_Y
)

@@

<br>  
<br>  
<br>  
******

@@

`early` and `late` will be divided by a number close to the median days after sowing when signs of the disease were first observed.
A number close to the median should ensure roughly even numbers in each category.
```{r median onset}
PM_dat_Y %>%
   mutate(first_sign_days = lubridate::ymd(first_sign_disease) -
             lubridate::ymd(planting_date)) %>%
   pull(first_sign_days)%>%
   median()
```
We will choose 42 days as this is the average days to flowering in mungbean. 
In addition Kelly et. al. [@-Kelly2017a] noted that yield losses are higher in cultivar Berken when disease establishes before flowering.

Now lets define these two new variables, `first_sign_days` and `onset` in the data.

```{r define_onset}
PM_dat_Y <- PM_dat_Y %>%
   mutate(first_sign_days = lubridate::ymd(first_sign_disease) -
             lubridate::ymd(planting_date)) %>%
   mutate(onset = case_when(first_sign_days <= 42 ~ "early_onset",
                            first_sign_days > 42 ~ "late_onset"))
```

Inspect a histogram of when first sign occurred in the crop
```{r hist_onset}
PM_dat_Y %>%
   ggplot(aes(x = as.integer(first_sign_days))) +
   geom_histogram(binwidth = 1) +
   geom_vline(xintercept = 42)
```

@@

```{r onset_Metafor-analysis}
PM_mv_onset <- rma.mv(
   yi = grain_yield.t.ha.,
   vi,
   mods = ~ spray_management * onset,
   method = "ML",
   random = list(~ spray_management | trial, ~ 1 | id),
   struct = "UN",
   control = list(optimizer = "optim"),
   data = PM_dat_Y
)

```{r}
anova(PM_mv, PM_mv_onset)$pval
```
This comparison tells us the meta-analysis including `onset` explained more residual heterogeneity, as shown by the lower QE value.
However the higher AIC indicates that the model omitting `onset` is a slightly better fit.
Also the chi-squared test uncertain that the `onset` model is significantly better as indicated by the `pval` `r round(anova(PM_mv, PM_mv_onset)$pval,3)`.
The higher log likelihood on the `onset` model does intricate there is some merit to this model and that perhaps with more trials to increase statistical power might provide a useful model.

@@

<br>  
<br>  
<br>  
******

@@

`high_pressure` and `low_pressure` will be divided by the median disease severity in the no spray control treatments at the end of the season.
```{r median_pressure}
PM_dat_Y %>%
   filter(fungicide_ai == "control") %>%
   pull(PM_final_severity)%>%
   median()
```
Now lets define these two new variables, `first_sign_days` and `onset` in the data.

```{r define_disease_pressure}
PM_dat_Y <- PM_dat_Y %>%
   mutate(d_pressure = case_when(PM_final_severity < 8 ~ "low_pressure",
                            PM_final_severity >= 8 ~ "high_pressure"))
```

Inspect a histogram of when first sign occurred in the crop
```{r hist_pressure}
PM_dat_Y %>%
   ggplot(aes(x = PM_final_severity)) +
   geom_histogram(binwidth = 1) +
   geom_vline(xintercept = 8)
```

<br>  
<br>  
<br>  
******

@@

```{r Dpressure_Metafor-analysis}
PM_mv_dp <- rma.mv(
   yi = grain_yield.t.ha.,
   vi,
   mods = ~ spray_management * d_pressure,
   method = "ML",
   random = list(~ spray_management | trial, ~ 1 | id),
   struct = "UN",
   control = list(optimizer = "optim"),
   data = PM_dat_Y
)

@@

```{r}
PM_dat_Y %>%
   group_by(spray_management, d_pressure)%>%
   summarise(n())
```

@@

<br>  
<br>  
<br>  
******

### Test variance-covavriance matrix structure
Let's try the model with an implied simpler variance-covariance matrix "compound symmetry" ("CS") to examine if unstructured matrix is suitable.
```{r}
PM_mv_cs <- rma.mv(
   yi = grain_yield.t.ha.,
   vi,
   mods = ~ spray_management,
   method = "ML",
   random = list(~ spray_management | trial, ~ 1 | id),
   struct = "CS",
   control = list(optimizer = "optim"),
   data = PM_dat_Y
)

@@

 > Therfore the model `PM_mv`, is a best fit.

@@

<br>
<br>
<br>
******

## Meta-analysis results summary
### Moderator estimates table
Let's present the meta-analysis results for the moderator variables in a table of estimates.

@@

yield_estimates_table <- 
   flextable(results_mv[c(2,5,6,3,4), c(1,3:8,10)]) %>%
   align(j = 3:8, align = "center", part = "all") %>%
   fontsize(size = 8, part = "body") %>%
   fontsize(size = 10, part = "header") %>%
   italic(italic = TRUE, part = "header") %>%
   set_caption("Table 2: Estimated mungbean yield mean difference to the no spray control (intercept) for each spray schedule treatment. Yield estimates (u) were calculated from a network meta-analysis of data obtained from grey literature reports of 'k' field trials undertaken in Eastern Australia. P values indicate statistical significance in comparison to the intercept.") %>%
   autofit() %>%
   footnote(i = 1,
            j = c(2:4,6:8),
            value = as_paragraph(c(
               "number of treatment means categorised to each spray schedule",
               "number of trials with the respective spray schedule",
               "estimated mean yield determined by the meta-analysis",
               "Lower range of the 95% confidence interval",
               "Upper range of the 95% confidence interval",
               "indicates the significance between each respective spray schedule and the no spray control (intercept)"
                                   )),
            ref_symbols = letters[1:6],
            part = "header", inline = TRUE
            )

@@

Save table to a word document to use in manuscript
```{r eval=FALSE, include=FALSE}
# Save table as a word document
doc <- officer::read_docx()
doc <- body_add_flextable(doc, value = yield_estimates_table)
print(doc, target = "paper/figures/Table2_yield_estimates.docx")
```

@@

### Moderator contrasts
Let's investigate the treatment contrasts

@@

These contrasts can be viewed in a plot
```{r plotContrasts_yield}
par(mar = c(5, 13, 4, 2) + 0.1)
plot(glht(PM_mv, linfct = cbind(contrMat(rep(
   1, 6
), type = "Tukey"))), yaxt = 'n')
axis(
   2,
   at = seq_along(contrast_Ssum$contrast),
   labels = rev(contrast_Ssum$contrast),
   las = 2,
   cex.axis = 0.8
)
```

@@

```{r netmeta-forest}
forest(
   net_con,
   reference.group = 1,
   rightcols = c("effect", "ci", "Pscore"),
   rightlabs = "P-Score",
   small.values = "bad"
)
```
<br>  
<br>  
<br>  
******

>>>>

```{r Yield_ME_libraries, echo=TRUE,results='hide'}
if (!require("pacman"))
   install.packages("pacman")
pacman::p_load(
   tidyverse,
   kableExtra,
   bomrang,
   RColorBrewer,
   metafor,
   here,
   netmeta,
   multcomp,
   flextable
)

@@

# Data
PM_dat_Y <-
   read.csv("cache/PM_yield_clean_data.csv", stringsAsFactors = FALSE
   )
```

@@

## Spray schedule meta-analysis

### Spray schedule moderator

@@

```{r Yield_Metafor-analysis}
PM_mv <- rma.mv(
   yi = grain_yield.t.ha.,
   V = vi,
   mods = ~ spray_management,
   method = "ML",
   random = list( ~ spray_management | trial, ~ 1 | id),
   struct = "UN",
   control = list(optimizer = "optim"),
   data = PM_dat_Y
)

@@

******

@@

`early` and `late` will be divided by a number close to the median days after sowing when signs of the disease were first observed.
A number close to the median should ensure roughly even numbers in each category.

```{r median onset}
PM_dat_Y %>%
   mutate(first_sign_days = lubridate::ymd(first_sign_disease) -
             lubridate::ymd(planting_date)) %>%
   pull(first_sign_days) %>%
   median()
```

We will choose 42 days as this is the average days to flowering in mungbean. 
In addition, Kelly et. al. [-@Kelly2017a] noted that yield losses are higher in cultivar Berken when disease establishes before flowering.

Now let's define these two new variables, `first_sign_days` and `onset` in the data.

```{r define_onset}
PM_dat_Y <- PM_dat_Y %>%
   mutate(first_sign_days = lubridate::ymd(first_sign_disease) -
             lubridate::ymd(planting_date)) %>%
   mutate(onset = case_when(
      first_sign_days <= 42 ~ "early_onset",
      first_sign_days > 42 ~ "late_onset"
   ))
```

Inspect a histogram of when first sign occurred in the crop.

```{r hist_onset}
PM_dat_Y %>%
   ggplot(aes(x = as.integer(first_sign_days))) +
   geom_histogram(binwidth = 1) +
   geom_vline(xintercept = 42)
```

@@

```{r onset_Metafor-analysis}
PM_mv_onset <- rma.mv(
   yi = grain_yield.t.ha.,
   V = vi,
   mods = ~ spray_management * onset,
   method = "ML",
   random = list( ~ spray_management | trial, ~ 1 | id),
   struct = "UN",
   control = list(optimizer = "optim"),
   data = PM_dat_Y
)
```

```{r}
anova(PM_mv, PM_mv_onset)
```

@@

This comparison tells us the meta-analysis including `onset` explained more residual heterogeneity, as shown by the lower QE value.
However the higher AIC indicates that the model omitting `onset` is a slightly better fit.
Also the chi-squared test is uncertain that the `onset` model is significantly better as indicated by the ANOVA `pval` `r round(anova(PM_mv, PM_mv_onset)$pval,3)`.
The higher log likelihood on the `onset` model does indicate that there is some merit to this model and that perhaps with more trials to increase statistical power it might be a useful model.  
<br>  
<br>  
<br>

******

@@

`high_pressure` and `low_pressure` will be divided by the median disease severity in the no spray control treatments at the end of the season.

```{r median_pressure}
PM_dat_Y %>%
   filter(fungicide_ai == "control") %>%
   pull(PM_final_severity) %>%
   median()
```

Now let's define these two new variables, `low_pressure` and `high_pressure` in the data.

```{r define_disease_pressure}
PM_dat_Y <- PM_dat_Y %>%
   mutate(
      d_pressure = case_when(
         PM_final_severity < 8 ~ "low_pressure",
         PM_final_severity >= 8 ~ "high_pressure"
      )
   )
```

Inspect a histogram of the final severity values in the crop.

```{r hist_pressure}
PM_dat_Y %>%
   ggplot(aes(x = PM_final_severity)) +
   geom_histogram(binwidth = 1) +
   geom_vline(xintercept = 8)
```

******

@@

We are adding the `d_pressure` variable as a moderator to test and interaction with the `spray_management` variable as a moderator and an interactive term to the `trial` random variable.

```{r Dpressure_Metafor-analysis}
PM_mv_dp <- rma.mv(
   yi = grain_yield.t.ha.,
   V = vi,
   mods = ~ spray_management * d_pressure,
   method = "ML",
   random = list( ~ spray_management | trial, ~ 1 | id),
   struct = "UN",
   control = list(optimizer = "optim"),
   data = PM_dat_Y
)
```

@@

```{r}
PM_dat_Y %>%
   group_by(spray_management, d_pressure) %>%
   summarise(n())
```

@@

******

### Test variance-covavriance matrix structure

Let's try the model with an implied simpler variance-covariance matrix, "compound symmetry" (CS), to examine if unstructured matrix is suitable.

```{r}
PM_mv_cs <- rma.mv(
   yi = grain_yield.t.ha.,
   V = vi,
   mods = ~ spray_management,
   method = "ML",
   random = list( ~ spray_management | trial, ~ 1 | id),
   struct = "CS",
   control = list(optimizer = "optim"),
   data = PM_dat_Y
)

@@

 > Therefore the model `PM_mv`, is a best fit.

@@

******

## Meta-analysis results summary

### Moderator estimates table

Let's present the meta-analysis results for the moderator variables in a table of estimates.

@@

yield_estimates_table <-
   flextable(results_mv[c(2, 5, 6, 3, 4), c(1, 3:8, 10)]) %>%
   align(j = 3:8, align = "center", part = "all") %>%
   fontsize(size = 8, part = "body") %>%
   fontsize(size = 10, part = "header") %>%
   italic(italic = TRUE, part = "header") %>%
   set_caption(
      "Table 2: Estimated mungbean yield mean difference to the no spray control (intercept) for each spray schedule treatment. Yield estimates (u) were calculated from a network meta-analysis of data obtained from grey literature reports of 'k' field trials undertaken in Eastern Australia. P values indicate statistical significance in comparison to the intercept."
   ) %>%
   autofit() %>%
   footnote(
      i = 1,
      j = c(2:4, 6:8),
      value = as_paragraph(
         c(
            "number of treatment means categorised to each spray schedule",
            "number of trials with the respective spray schedule",
            "estimated mean yield determined by the meta-analysis",
            "Lower range of the 95% confidence interval",
            "Upper range of the 95% confidence interval",
            "indicates the significance between each respective spray schedule and the no spray control (intercept)"
         )
      ),
      ref_symbols = letters[1:6],
      part = "header",
      inline = TRUE
   )

@@

Save table to a word document to use in manuscript.

```{r eval=FALSE, include=FALSE}
# Save table as a word document
doc <- officer::read_docx()
doc <- body_add_flextable(doc, value = yield_estimates_table)
print(doc, target = "paper/figures/Table2_yield_estimates.docx")
```

@@

### Moderator contrasts

Let's investigate the treatment contrasts

@@

These contrasts can be viewed in a plot.

```{r plotContrasts_yield}
par(mar = c(5, 13, 4, 2) + 0.1)
plot(glht(PM_mv, linfct = cbind(contrMat(rep(
   1, 6
), type = "Tukey"))), yaxt = 'n')
axis(
   2,
   at = seq_along(contrast_Ssum$contrast),
   labels = rev(contrast_Ssum$contrast),
   las = 2,
   cex.axis = 0.8
)
```

@@

```{r netmeta-forest}
forest(
   net_con,
   reference.group = 1,
   rightcols = c("effect", "ci", "Pscore"),
   rightlabs = "P-Score",
   small.values = "bad"
)
```
<br>  
<br>  
<br>

******
